Quantum-Classical Phase Transition for Exciton-Polariton Josephson Junctions

### 2017-01-07 writing/phase_transition/phase_transition.tex
Mathieu spectrum and other minor changes

Hamiltonian (\ref{eq:hamiltonian}) represents extended model for internal Josephson junction effect with exciton polariton BEC trapped in double well potential.
It is implies existence of quantum tunneling of the particles between wells with effective (nonlinear) tunneling rate
%
\begin{equation}
G_{eff} = \dfrac{1}{2}(G+\Gamma N - C\hat{\psi}_1^\dag\hat{\psi}_2).
\label{eq:tunneling_rate}
\end{equation}
%
The contribution from relevant coefficients on can be estimated from variational ansatz
%
\begin{equation}
\Phi_{\pm} = A_{\pm} \Big[ \exp \Big( -\dfrac{(x - x_0)^2}{2 a^2} \Big) \pm \exp \Big( -\dfrac{(x + x_0)^2}{2 a^2} \Big) \Big]
\label{eq:two_modes_eq}
\end{equation}
%
where $a$ is width of condensate wavefunctions at each wells (we suppose them equal to each other), defined from normalization conditions.
In Fig. \ref{pic:gamma_pm_vs_g} the behavior of normalized coefficients $\dfrac{\gamma_{ij}}{g}$ $i,j \in \{+,-\}$ as a function of half of intra-well distance $x_0$ is shown.
It is important that at differences between them rapidly vanishes with increasing of distance $x_0$.
We can suppose $\Gamma = C = 0$ for $x_0 >> a$.
This limit exactly corresponds to familiar problem of two weakly linked condensates at zero temperature \underline{- see e.g. ~\cite{39} and cf. }.
%
\begin{figure}[ht]
\center{\includegraphics[width=0.65\linewidth]{pic/ypm_x0.png}}
\caption{Dependence of $\gamma_{ij} / g$ versus $x_0 / a$ \label{pic:gamma_pm_vs_g}}
\end{figure}
%

@@

Inserting (\ref{eq:transformation_of_phi}) into (\ref{eq:schrodinger}) we come to convenient Schr\"odinger equation
%
\begin{equation}
\dfrac{d^2\Psi}{dx^2} + (E - V(x))\Psi = 0
\label{eq:schrodinger_usual}
\end{equation}
%
for quantum particle with mass $m_{eff} = \dfrac{1}{2}$, normalized energy $E \equiv \dfrac{E}{\alpha}$ moving in the potential $V(x) = s^2 V_0(x)$ with
%
\begin{equation}
V_0(x) = \frac{ (\frac{1}{4} \Lambda^2 - \lambda (1 - \lambda)) \sn^2{x} - \Lambda \cn{x}}{\dn^2{x}}
\label{eq:potential}
\end{equation}
%

@@

In purely quantum limit the two mode system approaches by the state $\ket{\psi} \propto (\psi_1^\dag)^s (\psi_1^\dag)^s \ket{0}_1 \ket{0}_2$ with limiting value of the phase wavefunction $\Phi(\phi) \to (2\pi)^{-1/2}$.
In this regime we distinguish two limits.
First, we suppose that inequality $\Lambda \ll \lambda \ll 1$ is holds.
In this case from (\ref{eq:potential}) one can obtain
%
\begin{equation}
V_F(z) = -\lambda(1 - \lambda) \sd^2{z}
\label{eq:fock1_potential}
\end{equation}.
%

{\red Mathieu equaiton and it's spectrum}

>>>>

Hamiltonian (\ref{eq:hamiltonian}) represents extended model for internal Josephson junction effect with exciton polariton BEC trapped in double well potential.
It is implies existence of quantum tunneling of the particles between wells with effective (nonlinear) tunneling rate
%
\begin{equation}
G_{eff} = \dfrac{1}{2}(G+\Gamma N - C\hat{\psi}_1^\dag\hat{\psi}_2).
\label{eq:tunneling_rate}
\end{equation}
%
The contribution from relevant coefficients on can be estimated from variational ansatz
%
\begin{equation}
\Phi_{\pm} = A_{\pm} \Big[ \exp \Big( -\dfrac{(x - x_0)^2}{2 a^2} \Big) \pm \exp \Big( -\dfrac{(x + x_0)^2}{2 a^2} \Big) \Big]
\label{eq:two_modes_eq}
\end{equation}
%
where $a$ is width of condensate wavefunctions at each wells (we suppose them equal to each other), defined from normalization conditions.
In Fig. \ref{pic:gamma_pm_vs_g} the behavior of normalized coefficients $\dfrac{\gamma_{ij}}{g}$ $i,j \in \{+,-\}$ as a function of half of intra-well distance $x_0$ is shown.
It is important that at differences between them rapidly vanishes with increasing of distance $x_0$.
We can suppose $\Gamma = C = 0$ for $x_0 >> a$.
This limit exactly corresponds to familiar problem of two weakly linked condensates at zero temperature {\red see e.g. \cite{39} and cf. }.
%
\begin{figure}[ht]
\center{\includegraphics[width=0.65\linewidth]{pic/ypm_x0.png}}
\caption{Dependence of $\gamma_{ij} / g$ versus $x_0 / a$ \label{pic:gamma_pm_vs_g}}
\end{figure}
%

@@

Inserting (\ref{eq:transformation_of_phi}) into (\ref{eq:schrodinger}) we come to convenient Schr\"odinger equation
%
\begin{equation}
\frac{d^2\Psi}{dz^2} + (E - V(z))\Psi = 0
\label{eq:schrodinger_usual}
\end{equation}
%
for quantum particle with mass $m_{eff} = \dfrac{1}{2}$, normalized energy $E \equiv \dfrac{E}{\alpha}$ moving in the potential $V(z) = s^2 V_0(z)$ with
%
\begin{equation}
V_0(z) = \frac{ (\frac{1}{4} \Lambda^2 - \lambda (1 - \lambda)) \sn^2{z} - \Lambda \cn{z}}{\dn^2{z}}
\label{eq:potential}
\end{equation}
%

@@

In purely quantum limit the two mode system approaches by the state $\ket{\psi} \propto (\psi_1^\dag)^s (\psi_1^\dag)^s \ket{0}_1 \ket{0}_2$ with limiting value of the phase wavefunction $\Phi(\phi) \to (2\pi)^{-1/2}$.
In this regime we distinguish two limits.
First, we suppose that inequality $\Lambda \ll \lambda \ll 1$ is holds.
In this case from (\ref{eq:potential}) one can obtain
%
\begin{equation}
V_F(z) = -\lambda(1 - \lambda) \sd^2{z}.
\label{eq:fock1_potential}
\end{equation}.
%

{\red
Using the fact that $\lambda$ is small we can replace the potential (\ref{eq:fock1_potential}) with
%
\begin{equation}
V_F(z) \sim -\lambda (1 - \lambda) \sin^2{z},
\end{equation}
%
and our Schrodinger equation transforms to the so-called Mathieu equation
%
\begin{equation}
\frac{d^2 \Psi}{d z^2} + (a - 2q  \cos{2z}) \Psi = 0.
\end{equation}
%
where $a = E + s^2 \lambda(1 - \lambda) / 2$, $q = s^2 \lambda (1 - \lambda)/ 4$.
We recall our boundary condition: $\Psi(z + 2\pi) = \Psi(z)$.
So, spectrum consist of all Mathieu functions $ce_0(z, q)$, $ce_1(z, q)$, $se_1(z, q)$, \dots, which are $2\pi$-periodic, and corresponding eigenvalues $a$ can be found using series expansions in powers of $q$ (see Abramowitz and Stegun).
}
